3.9: Logarithmic Regression

### 2019-11-22 static/slides/21-slides.Rmd
Minor formatting tweaks to slides 21
Add comparing models table to slides 21

- A **$25K fall in GDP** $\rightarrow$ a $(-25 \times 1.3\%)=$ **32.5% decrease** in Life Expectancy

- A **$100K rise in GDP** $\rightarrow$ a $(100 \times 1.3\%)=$ **130% increase** in Life Expectancy

@@

- A **$1% change in GDP** $\rightarrow$ a **0.147% increase** in Life Expectancy

@@

| Model | Equation | Interpretation |
|-------|----------|----------------|
| Linear-Log | $Y=\beta_0+\beta_1 \mathbf{ln(X)}$ | 1% change in $X \rightarrow \frac{\hat{\beta_1}}{100}$ **unit** change in $Y$ |
| Log-Linear | $\mathbf{ln(Y)}=\beta_0+\beta_1X$ | 1 **unit** change in $X \rightarrow \hat{\beta_1}\times 100$**%** change in $Y$ |
| Log-Log | $\mathbf{ln(Y)}=\beta_0+\beta_1\mathbf{ln(X)}$ | 1**%** change in $X \rightarrow \hat{\beta_1}$**%** change in $Y$ |

- Hint: the variable that gets logged changes in **percent** terms, the variable not logged changes in **unit** terms

@@

@@

- In practice, the following types of variables are logged (ln):
  - Variables that must always be positive (prices, sales, market values)
  - Very large numbers (population, GDP)
  - Variables expressed as percentages or percentage changes (inflation, population growth, GDP growth, labor force participation rate, unemployment rate)
  - Variables that have nonlinear scatterplots

@@

- Never use logs for:
  - Variables that are less than one, decimals, 0, or negative (temperature)
  - Categorical variables
  - Time variables (year, week, day)

>>>>

- A **$100 (thousand) rise in GDP** $\rightarrow$ a $(100 \times 1.3\%)=$ **130% increase** in Life Expectancy

@@

- A **1% change in GDP** $\rightarrow$ a **0.147% increase** in Life Expectancy

@@

| Model | Equation | Interpretation |
|-------|----------|----------------|
| Linear-**Log** | $Y=\beta_0+\beta_1 \mathbf{ln(X)}$ | 1% change in $X \rightarrow \frac{\hat{\beta_1}}{100}$ **unit** change in $Y$ |
| **Log**-Linear | $\mathbf{ln(Y)}=\beta_0+\beta_1X$ | 1 **unit** change in $X \rightarrow \hat{\beta_1}\times 100$**%** change in $Y$ |
| **Log**-**Log** | $\mathbf{ln(Y)}=\beta_0+\beta_1\mathbf{ln(X)}$ | 1**%** change in $X \rightarrow \hat{\beta_1}$**%** change in $Y$ |

- Hint: the variable that gets logged changes in **percent** terms, the variable not logged changes in *unit* terms

@@

# Comparing Models III

| Linear-Log | Log-Linear | Log-Log |
|:----------:|:----------:|:-------:|
| ![](21-slides_files/figure-html/unnamed-chunk-14-1.png) | ![](21-slides_files/figure-html/unnamed-chunk-20-1.png) | ![](21-slides_files/figure-html/unnamed-chunk-25-1.png) |
| $\hat{Y_i}=\hat{\beta_0}+\hat{\beta_1}\mathbf{ln(X_i)}$ | $\mathbf{ln(\hat{Y_i})}=\hat{\beta_0}+\hat{\beta_1}X_i$ | $\mathbf{ln(\hat{Y_i})}=\hat{\beta_0}+\hat{\beta_1}\mathbf{ln(X_i)}$ |
| $R^2=0.65$ | $R^2=0.30$ | $R^2=0.61$ |

@@

- In practice, the following types of variables are logged:
  - Variables that must always be positive (prices, sales, market values)
  - Very large numbers (population, GDP)
  - Variables we want to talk about as percentage changes or growth rates (money supply, population, GDP)
  - Variables that have diminishing returns (output, utility)
  - Variables that have nonlinear scatterplots

@@

- Avoid logs for:
  - Variables that are less than one, decimals, 0, or negative
  - Categorical variables (season, gender, political party)
  - Time variables (year, week, day)
